---
title: "Statistical Learning"
subtitle: "Chapter 3 - Exercise 10"
output: 
  html_document:
    # css: style_notebook.css
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
    toc_depth: 3
    number_sections: false
    theme: flatly
    highlight: tango
    code_folding: hide
    mathjax: "default"
    self_contained: true
    keep_md: false
---

```{css, echo=FALSE}
body {
  background-color: #f5f5f5;
}

.title {
  color: #4292c6;
  display: flex;
  justify-content: center;
  align-items: center;
}

.proof {
  background-color: #f5f5f5;
  padding: 5px;
  margin-bottom: 0;
  width: 100%;
  overflow: auto;
}

.special {
  display: inline-block;
  font-size: 30px;
  background-color: #e5f5e0;
  padding: 4px;
  margin-top: 2px;
  margin-bottom: 2px;
  width: fit-content;
  border-left: 3px solid #73AD21;
  transition: all ease-in-out 2s;
  animation: item-hover-off .5s;
}

.special:hover {
    transform: scale(1.5);
    transition: all 1s ease;
}

.row {
  text-align: center;
}

.special-text{
  display: inline-block;
  text-align: center;
}

.selector{
  width: 50%;
  float: left;
}

blockquote {
  background: #f9f9f9;
  border-left-color: #1ca2c3;
}
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  comment = ""
)
library(magrittr)
```

> This questions should be answered using the Carseats data set.

```{r}
car_seats <- ISLR::Carseats
```

```{r}
SmartEDA::ExpData(car_seats, type = 1)
```

```{r}
SmartEDA::ExpData(car_seats, type = 2)
```

```{r}
base::table(car_seats$Education)
```

## Part a

> Fit a multiple regression model to predict Sales using Price, Urban, and US.

```{r}
model1_fit <- stats::lm(Sales ~ Price + Urban + US, data = car_seats)
```

```{r}
base::summary(model1_fit)
```

## Part b

> Provide an interpretation of each coefficient in the model.

```{r}
var_price <- if(stats::coef(model1_fit)[["Price"]] < 0) "drop" else "raise"
```

- `Price` = `r stats::coef(model1_fit)[["Price"]]` indicates that sales will `r var_price`, on average, by `r base::abs(base::round(stats::coef(model1_fit)[["Price"]] * 1000))` units for every one dollar increase in their price (when holding `Urban` and `US` constant).

```{r}
var_urbanyes <- if(stats::coef(model1_fit)[["UrbanYes"]] < 0) "lower" else "higher"
```

- `UrbanYes` = `r stats::coef(model1_fit)[["UrbanYes"]]` indicates that sales are, on average, `r base::abs(base::round(stats::coef(model1_fit)[["UrbanYes"]] * 1000))` units `r var_urbanyes` for stores in urban areas, compared to stores in non-Urban areas (when holding `Price` and `US` constant).

```{r}
var_usyes <- if(stats::coef(model1_fit)[["USYes"]] < 0) "lower" else "higher"
```

- `USYes` = `r stats::coef(model1_fit)[["USYes"]]`indicates that sales are, on average, `r base::abs(base::round(stats::coef(model1_fit)[["USYes"]] * 1000))` units `r var_usyes` for stores in urban areas, compared to stores in non-Urban areas (when holding `Price` and `Urban` constant).

## Part c

> Write out the model in equation form, being careful to handle the qualitative variables properly.

$$Sales = 13.04 - 0.05 * Price - 0.02 * \text{I}_{(Urban = Yes)} + 1.2 * \text{I}_{(US = Yes)}$$

## Part d

> For which of the predictors can you reject the null hypothesis $H_0: \beta_j = 0$

Based on *p-vales* for `Price` and `US`.

## Part e

> On the basis of your response to the previous question, fit a smaller model that only uses the predictors for which there is evidence of association with the outcome.

```{r}
model2_fit <- stats::lm(Sales ~ Price + US, data = car_seats)
```

```{r}
base::summary(model2_fit)
```

## Part f

> How well do the models in (a) and (e) fit the data?

Based on $RSE$ and $R^2$, both models provide a good fit. Yet, the model from part e is preferable since it has a slightly lower $RSE$ (`r base::round(base::summary(model2_fit)$sigma, 3)` for model e Vs. `r base::round(base::summary(model1_fit)$sigma, 3)` for model a), and one less predictor.

## Part g

> Using the model from (e), obtain 95% confidence intervals for the coefficient(s).

```{r}
stats::confint(model2_fit)
```

## Part h

> Is there evidence of outliers or high leverage observations in the model from (e)?

```{r}
model2_fit_diagnose <- lindia::gg_diagnose(
  fitted.lm = model2_fit,
  plot.all = FALSE
)
```

```{r, fig.align='center', out.width='100%'}
model2_fit_diagnose$res_fitted +
  ggplot2::geom_smooth(
    se = FALSE
  ) +
  ggplot2::theme(
    plot.background = ggplot2::element_rect(
      fill = "#F5F5F5",
      color = "#F5F5F5"
    )
  )
```

```{r, fig.align='center', out.width='100%'}
model2_fit_diagnose$scalelocation +
  ggplot2::labs(
    x = "Fitted Values",
    y = "sqrt(std res)"
  ) +
  ggplot2::theme(
    plot.background = ggplot2::element_rect(
      fill = "#F5F5F5",
      color = "#F5F5F5"
    )
  )
```

```{r, fig.align='center', out.width='100%'}
model2_fit_diagnose$qqplot +
  ggplot2::theme(
    plot.background = ggplot2::element_rect(
      fill = "#F5F5F5",
      color = "#F5F5F5"
    )
  )
```

```{r, fig.align='center', out.width='100%'}
model2_fit_diagnose$cooksd +
  ggplot2::theme(
    plot.background = ggplot2::element_rect(
      fill = "#F5F5F5",
      color = "#F5F5F5"
    )
  )
```

```{r, fig.align='center', out.width='100%'}
model2_fit_diagnose$resleverage +
  ggplot2::theme(
    plot.background = ggplot2::element_rect(
      fill = "#F5F5F5",
      color = "#F5F5F5"
    )
  )
```
