---
title: "Linear Regression"
subtitle: "Chapter 3 - Exercise 8"
output: 
  html_document:
    # css: style_notebook.css
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
    toc_depth: 3
    number_sections: false
    theme: flatly
    highlight: tango
    code_folding: hide
    mathjax: "default"
    self_contained: true
    keep_md: false
---

```{css, echo=FALSE}
body {
  background-color: #f5f5f5;
}

.title {
  color: #4292c6;
  display: flex;
  justify-content: center;
  align-items: center;
}

.proof {
  background-color: #f5f5f5;
  padding: 5px;
  margin-bottom: 0;
  width: 100%;
  overflow: auto;
}

.special {
  display: inline-block;
  font-size: 30px;
  background-color: #e5f5e0;
  padding: 4px;
  margin-top: 2px;
  margin-bottom: 2px;
  width: fit-content;
  border-left: 3px solid #73AD21;
  transition: all ease-in-out 2s;
  animation: item-hover-off .5s;
}

.special:hover {
    transform: scale(1.5);
    transition: all 1s ease;
}

.row {
  text-align: center;
}

.special-text{
  display: inline-block;
  text-align: center;
}

.selector{
  width: 50%;
  float: left;
}

blockquote {
  background: #f9f9f9;
  border-left-color: #1ca2c3;
}
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  comment = ""
)
library(magrittr)
```

> This question involves the use of simple linear regression on the `Auto` data set.

```{r}
auto <- ISLR::Auto
```

## Part a

> Use the `lm` function to perform a simple linear regression with mpg as the response and horsepower as the predictor. Use the `summary` function to print the results. Comment on the output.

```{r}
auto_fit <- stats::lm(mpg ~ horsepower, data = auto)
```

```{r}
base::summary(auto_fit)
```

> i. Is there a relationship between the predictor and the response?

Yes, since the coefficient is statistically significant

> ii. How strong is the relationship between the predictor and the response?

Very much so since de *p-value* is very close to zero

> iii. Is the relationship between the predictor and the response positive or negative?

$\hat{\beta_1} < 0$, therefore the relationship is negative

> iv. What is the predicted mpg associated with a horsepower of 98? What are the associated 95% confidence and prediction intervals?

```{r}
stats::predict(auto_fit, tibble::tibble(horsepower = base::c(98)))
```

```{r}
stats::predict(auto_fit, tibble::tibble(horsepower = base::c(98)), interval = "confidence")
```

## Part b

```{r, fig.align='center', out.width='100%'}
ggplot2::ggplot(
  data = auto,
  mapping = ggplot2::aes(
    x = horsepower,
    y = mpg
  )
) + 
  ggplot2::geom_point() +
  ggplot2::geom_smooth(
    method = "lm",
    se = FALSE
  ) +
  ggplot2::theme(
    plot.background = ggplot2::element_rect(
      fill = "#F5F5F5",
      color = "#F5F5F5"
    )
  )
```

## Part c

```{r}
auto_fit_diagnose <- lindia::gg_diagnose(
  fitted.lm = auto_fit,
  plot.all = FALSE
)
```

```{r, fig.align='center', out.width='100%'}
auto_fit_diagnose$res_fitted +
  ggplot2::geom_smooth(
    se = FALSE
  ) +
  ggplot2::theme(
    plot.background = ggplot2::element_rect(
      fill = "#F5F5F5",
      color = "#F5F5F5"
    )
  )
```

```{r, fig.align='center', out.width='100%'}
auto_fit_diagnose$scalelocation +
  ggplot2::labs(
    x = "Fitted Values",
    y = "sqrt(std res)"
  ) +
  ggplot2::theme(
    plot.background = ggplot2::element_rect(
      fill = "#F5F5F5",
      color = "#F5F5F5"
    )
  )
```

```{r, fig.align='center', out.width='100%'}
auto_fit_diagnose$qqplot +
  ggplot2::theme(
    plot.background = ggplot2::element_rect(
      fill = "#F5F5F5",
      color = "#F5F5F5"
    )
  )
```

```{r, fig.align='center', out.width='100%'}
auto_fit_diagnose$cooksd +
  ggplot2::theme(
    plot.background = ggplot2::element_rect(
      fill = "#F5F5F5",
      color = "#F5F5F5"
    )
  )
```

```{r, fig.align='center', out.width='100%'}
auto_fit_diagnose$resleverage +
  ggplot2::theme(
    plot.background = ggplot2::element_rect(
      fill = "#F5F5F5",
      color = "#F5F5F5"
    )
  )
```

